
---
title: Activity analysis
author: Feng Gao
date: March 30, 2021
output: html_document
---



## Loading and preprocessing the data
 
```{r, echo=T, eval=T} 

  # readin data
  setwd("C:\\Users\\fgao1\\OneDrive\\Desktop\\scc_training_2021\\")
  getwd()
  gao1<-read.csv("activity.csv", header=T)
  names(gao1)
 
 ```

## What is average number of steps taken per day?

 histogram of mean is shown in Figure 1 and the median daily steps is listed in the table
 
 ```{r, echo=T, eval=T} 

  # total steps by days
  date.list<-unique(gao1$date)
  n.date<- length(date.list)
  mean.steps<-rep(NA, n.date)
  median.steps<-rep(NA, n.date)

  for (i in 1:n.date) {
     date.i<-date.list[i]
     step.i<-gao1[gao1$date==date.i, "steps"]
     mean.steps[i]<- mean(step.i, na.rm=T)
     median.steps[i]<- quantile(step.i, probs=0.5, na.rm=T)
  }
  median.steps
  
  hist(mean.steps, main="Figure 1: mean daily steps")

  
 ```
 
## What is the average daily activity pattern?
  time series plot of the average number of steps 
  (averaged across all days) is shown in Figure 2

 ```{r, echo=T, eval=T} 

   # mean steps by interval
  int.list<-unique(gao1$interval)
  n.int<- length(int.list)
  mean.steps<- array(NA, c(n.int,2))

  for (i in 1:n.int) {
     int.i<-int.list[i]
     step.i<-gao1[gao1$interval==int.i, "steps"]
     mean.steps[i,1]<-int.i
     mean.steps[i,2]<-mean(step.i, na.rm=T)
  }
  
  plot(x=mean.steps[,1], y=mean.steps[,2], type="l",
       main="Figure 2: mean daily pattern", xlab="time intervals (5 mins)")
  
  
```
# Which 5-minute interval contain the maximum number of steps?

```{r, echo=T, eval=T} 

max.steps.loc <- mean.steps[which.max(mean.steps[,2]),1]
max.steps.loc

```





## Imputing missing values
 a new variable steps.new will be created where the missing values 
 will be replaced by the overall average

```{r, echo=T, eval=T} 

gao1$steps.new <- ifelse(is.na(gao1$steps), mean(gao1$steps, na.rm=T), gao1$steps)
cat("\n     raw-data: mean =", round(mean(gao1$steps, na.rm=T), 2), ",   SD =", round(sqrt(var(gao1$steps, na.rm=T)), 2))
cat("\n imputed-data: mean =", round(mean(gao1$steps.new), 2), ",   SD =", round(sqrt(var(gao1$steps.new)), 2))

```

## Are there differences in activity patterns between weekdays and weekends?
 panel plots for weekdays and weekends patterns (figure3)
 
```{r, echo=T, eval=T} 
library(lattice)
gao1$date.new<- weekdays(as.Date(gao1$date))
gao1$weekend<- ifelse(gao1$date.new == "Saturday" | gao1$date.new == "Sunday", "WeekEnd", "WeekDay")
mean.weekend<- aggregate(steps ~ interval+weekend, data=gao1, FUN=mean)
xyplot(steps ~ interval | factor(weekend), layout = c(1, 2), type = "l",
               xlab = "time intervals (5 min)", ylab = "Average Number of Steps",  
               data = mean.weekend)
```


